---
title: "Writing Functions"
author: "Emma Warshaw"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)
```

```{r}
set.seed(1)
```

##Do something simple

My first function
The best way to build up a function is to start with code you’ve written outside a function. To see how this might work, I’ll start with a simple example: the chunk below takes a sample from a normal distribution and then computes the vector of Z scores for the sample.

Set the function, then take the x-vector, subtract the mean and divide by the standard deviation.  This gives you a z-score and is the type of thing you would want to do in a function. 
```{r}
x_vec = rnorm(30, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)
```
If I want to repeat this (admittedly simple) process for lots of samples, I might want to have a function that takes the sample as an argument, computes the vector of Z scores in the body, and returns the result. I define such a function below.

Body of the function exists between the curly brackets. Inside the function, you want to use the same thing everywhyere that you want to operate on the input.  In this case, it's x. 

Then, use the functions z_scores to operate on the x_vec that we created earlier. 

```{r}
z_scores = function(x) {
  if (!is.numeric(x)) {
    stop("Input must be numeric")
  }
  if (length(x) < 3){
    stop("Intput must have at least 3 numbers")
  }
  z = (x - mean(x)) / sd(x)
  z
}

z_scores(x_vec)
```

I can try this with a few samples and confirm that it works. I should also try to think of ways this code might break; the attempts below try a variety of inputs to see what happens.

The final line works, but we'd really prefer it to break.  Z-scores for these character variables are nonsensical but it's coercing into numeric because it senses a pattern. We're going to edit the function to account for this. 

***These should give errors***
```{r, error = TRUE}
z_scores(3)
z_scores("my name is jeff")
z_scores(mtcars)
z_scores(c(TRUE, TRUE, FALSE, TRUE))
```

These all did something I didn’t want, but only two returned errors. To avoid behavior you don’t want (i.e. to “fail noisily and as soon as possible”) we’ll add some checks on the argument values using conditional statements.
```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = mean(x) / sd(x)
  
  z
}
```
Fantastic – we have a pretty solid function for computing Z scores!

## Multiple outputs

In some cases it might be better to return the mean and standard deviation instead of the Z scores. A first option is to store each of the values in a named list, and to return that list. (We’ll talk more about lists in iteration and listcols.)

This will spit two numbers out at the same time. 

```{r}
mean_and_sd = function(x) {
  if (!is.numeric(x)) {
    stop("Input must be numeric")
  }
  if (length(x) < 3){
    stop("Intput must have at least 3 numbers")
  }
  mean_x = mean(x)
  sd_x = sd(x)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```
Check that the function works.  Create a new vector for fun. 
```{r}
xvec = rnorm(1000)
mean_and_sd(x_vec)
```

Alternatively, we might store values in a data frame.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```
In general, either of these will be fine; which one you choose will depend on what kind of values you want to return, and what you plan to do with the function itself. If you want to return the original sample along with the computed values, a list might make sense; if you plan to run your function a lot and study the results, having a data frame will make it easier to use other tools. We’ll see more of that in iteration and simulation.

## Multiple inputs

As exciting as it is to compute Z scores, let’s start setting our sights higher. I’d like to have a function that takes a given sample size along with a true mean and standard deviation, simulates data from a normal distribution, and returns the estimated mean and standard deviation. I’ll start from the code below.

Little more complex: creating a tibble then funding some summary stats for the tibble.  

```{r}
sim_data = tibble(
  x = rnorm(n = 100, mean = 4, sd = 3)
)

sim_data %>% 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
  )
```

You should take a few minutes to examine this code – make a plot of the simulated data to make sure it “makes sense”, take a look at the result of computing the mean and standard deviation, etc.

Once you’re satisfied, it’s time to wrap things up in a function. I’d like to be able to change the sample size and parameters, so those will be my arguments; the code that simulates data and computes the sample mean and SD go in the body; and the return statement should include the estimates. A function that does all this, using default values for the mean and standard deviation, is below.

```{r}
sim_mean_sd = function(n, mu = 2, sigma = 3) {
  
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  sim_data %>% 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )
}

sim_mean_sd(100, 6, 3)
sim_mean_sd(n = 100, mu = 6, sigma = 3) 
```
Repeated calls to sim_mean_sd() will give a sense of sampling variability in estimating the mean and standard deviation from a sample; take a few minutes to run sim_mean_sd(30) a few times, and then to run sim_mean_sd(300), and think about the results. We’ll examine this more formally in iteration and in simulation.

This is also a good time to point out how R handles argument matching. We can use positional matching, meaning the first value supplied is taken to be the first argument, the second value is the second argument, and so on. We do this with tidyverse functions a lot; the first argument is always a dataframe, and we just supply that dataframe in the first position. We also use positional matching when we call mean(x) or sim_mean_sd(30, 5, 1).

Alternatively, we can use named matching, which uses the argument name in the function call. Named matching can be a bit more stable when you’re writing your own functions (in case you decide to change the order of the inputs, for example) but isn’t strictly necessary. Named arguments can be supplied in any order: sim_mean_sd(n = 30, mu = 5, sd = 1) is equivalent to sim_mean_sd(sd = 1, n = 30, mu = 5).

Revisiting past examples
There have been a couple of times in this class that I’ve had to write code I didn’t like, because it would have made sense to write a function. We’ll revisit those quickly to see how we could improve our code.

## Let's review Napoleon Dynamite

Scraping Amazon

In reading data from the web, we wrote code that allowed us to scrape information in Amazon reviews. That code is below.

```{r}
url = "https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber=1"

dynamite_html = read_html(url)

review_titles = 
  dynamite_html %>%
  html_nodes(".a-text-bold span") %>%
  html_text()

review_stars = 
  dynamite_html %>%
  html_nodes("#cm_cr-review_list .review-rating") %>%
  html_text() %>%
  str_extract("^\\d") %>%
  as.numeric()

review_text = 
  dynamite_html %>%
  html_nodes(".review-text-content span") %>%
  html_text() %>% 
  str_replace_all("\n", "") %>% 
  str_trim()

reviews = tibble(
  title = review_titles,
  stars = review_stars,
  text = review_text
)
```
This takes the first page of reviewsm but we wat all the reviews.  Let's write a function to scrape all the pages of reviews that we want. 

Let’s write a quick function to scrape review information for any URL to an Amazon review page.
```{r}
read_page_reviews <- function(url) {
  
  dynamite_html = read_html(url)
  
  review_titles = 
    html %>%
    html_nodes(".a-text-bold span") %>%
    html_text()
  
  review_stars = 
    html %>%
    html_nodes("#cm_cr-review_list .review-rating") %>%
    html_text() %>%
    str_extract("^\\d") %>%
    as.numeric()
  
  review_text = 
    html %>%
    html_nodes(".review-text-content span") %>%
    html_text() %>% 
    str_replace_all("\n", "") %>% 
    str_trim() %>% 
    str_subset("The media could not be loaded.", negate = TRUE) %>% 
    str_subset("^$", negate = TRUE)
  
  reviews = 
    tibble(
    title = review_titles,
    stars = review_stars,
    text = review_text
  )
  
  reviews
}
```
Next we’ll use this to read in reviews from a few pages and combine the results.
```{r}
url_base = "https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber="
vec_urls = str_c(url_base, 1:5)

dynamite_reviews = bind_rows(
  read_page_reviews(vec_urls[1]),
  read_page_reviews(vec_urls[2]),
  read_page_reviews(vec_urls[3]),
  read_page_reviews(vec_urls[4]),
  read_page_reviews(vec_urls[5])
)

dynamite_reviews
```
Loading LoTR data
In tidy data, we broke the “only copy code twice” rule when we used the code below to process the LoTR words data:

```{r}
fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  gather(key = sex, value = words, female:male) %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 
```

Functions as arguments
One powerful tool is the ability to pass functions as arguments into functions. This might seem like a weird thing to do, but it has a lot of handy applications – we’ll see just how far it goes in the next modules in this topic.

As a quick example, suppose we wanted to get a sense of how similar or different values in a vector are to each other. There are lots of ways to measure this – variance, standard deviation, range, inter-quartile range – and some are more appropriate in some cases than in others. The function below allows you to input a vector and a function, and returns the result of applying the specified function to the vector input.

```{r}
x_vec = rnorm(25, 0, 1)

my_summary = function(x, summ_func) {
  summ_func(x)
}

my_summary(x_vec, sd)

my_summary(x_vec, IQR)

my_summary(x_vec, var)
```

This example is pretty trivial – you could just apply those functions directly to x and skip the hassle – but in many cases the idea of passing functions as arguments is really powerful. As a practical example, remember that you can reorder factors according to different summaries in fct_reorder!

Scoping and names
Take a look at the code below. Will the call f(x = y) work? If so, what will it produce? What is the current value of x, y, and z?

```{r}
f = function(x) {
  z = x + y
  z
}

x = 1
y = 2

f(x = y)
```

Examples like this are tricky, but emphasize an issue that comes up a lot in writing functions: you define a variable in your global environment and use it in your function, but it isn’t passed as an argument. This is easy to miss, especially when you go from code written in chunks to a function, and can be hard to track down if you empty your working directory or change a variable name. The best advice I have is to give your arguments useful names and think carefully about where everything is defined, and to periodically restart R and try everything again!


